Smart Drill-Down

### 2015-12-25 TableSummarization/Paper/algorithms.tex
addressed most reviewer comments, brought back to length

\begin{lemma}\label{lemma:rule-ordering}
Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
$\text{Score}(R^{\prime}) \geq \text{Score}(R)$.
\end{lemma}
\papertext{The proof of this lemma, as well as other proofs, can be found in the appendix of the technical report~\cite{tr}}\techreporttext{The proof of this lemma, as well as other proofs, can be found in the appendix}. 
Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:

@@

\stitle{High-Level Procedure:} Now, based on the submodularity property, the greedy procedure,
shown in Algorithm~\ref{algo:best-rule-set}, has desirable approximation guarantees. Since Score is a submodular function of the set $S$, 
this greedy procedure is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.

The expensive step in the above procedure is the step where the Score is computed for every
single rule. Given the number of rules can be large, this can be an especially time-consuming process.

@@

The parameter $m_w$ stands for \textit{Max Weight}. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by BRS does not contain $r$. BRS runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, its Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$). Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small. \papertext{In the technical report, we describe strategies for setting $m_w$ as well as other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}

@@

\techreporttext{
\stitle{Setting parameters $W$, $k$, $m_w$:} \mrj{Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.}

\mrj{Parameter $k$ is the number of new rules to display upon each smart drill-down. Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information. On the other hand, small values of $k$ may display too little information about the table. Fortunately, the BRS algorithm is incremental in nature. That is, in order to find the best rule list of size $k+1$, it first finds the best rule-list of size $k$, and then finds another rule to add to get a rule-list of size $k+1$. Thus instead of running the algorithm with a fixed value of $k$, it can start with an empty rule-list and keep adding rules to it, displaying new rules as they are found. This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search. Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.}

\mrj{$W$ is the weight function that determines which rules are interesting. This is a function specified by the user as a black box. Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu. In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface. The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.}

\mrj{The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time. Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time. We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling. We create a small random sample of tuples from the table, and run the BRS algorithm on it. Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output. To account for sampling error, we can set $m_w$ to $2x$, which works well in practice.
}
}

>>>>

\begin{lemma}\label{lemma:rule-ordering}
Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
$\text{Score}(R^{\prime}) \geq \text{Score}(R)$.
\end{lemma}
\papertext{The proof of this lemma, as well as other proofs, can be found in the technical report~\cite{tr}}\techreporttext{
\begin{proof}
The score of rule list $R$ is given by $$\text{Score}(R) = \sum_{r \in R} W(r) \times MCount(r,R)$$
For each tuple $t$, let $TOP(t,R)$ denote the first rule in rule-list $R$ that covers $t$. Then the MCount of rule $r$ in $R$ is simply $\sum_{t \in T : TOP(t,R) = r} 1$. Thus Score can be rewritten as:
$$\text{Score}(R) = \sum_{t \in T} W(TOP(t,R))$$
where we set $W(TOP(t)) = 0$ when $t$ is not covered by any rule in $R$. Now say we have rule lists $R$, $R'$ which have the same rules, but where $R'$ has rules in decreasing order by weight. For any tuple $t$ covered by $R$, let $r'$ be the highest weight rule in $R$ that covers $t$. Let $r$ be the first rule in $R$ that covers $t$. Then we have $TOP(t,R) = r$, $TOP(t, R') = r'$ and $W(r') \geq W(r)$. This gives us $W(TOP(t,R')) \geq W(TOP(t,R))$. Adding these inequalities for all $t$ covered by $R$ gives us $\text{Score}(R') \geq \text{Score}(R)$ as required.
\end{proof}
}. 
Thus, it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a \emph{set} of rules as follows:

@@

\stitle{High-Level Procedure:} Based on the submodularity property, the greedy procedure,
shown in Algorithm~\ref{algo:best-rule-set}, has desirable approximation guarantees. Since Score is a submodular function of the set $S$, 
this greedy procedure is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum (actually, it is $1 - (\frac{k-1}{k})^k$ for $k$ rules, which is much better for small $k$).

The expensive step in the above procedure is the step where the Score is computed for every
single rule. Given the number of rules can be as large as the table itself, this can be especially time-consuming.

@@

The parameter $m_w$ stands for \textit{Max Weight}. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by BRS does not contain $r$. BRS runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, its Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$). Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small. \papertext{In the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}

@@

\techreporttext{
\stitle{Setting parameters $W$, $k$, $m_w$:} Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.

Parameter $k$ is the number of new rules to display upon each smart drill-down. Large values of $k$ increase the run-time quadratically, and can also overwhelm the user with too much information. On the other hand, small values of $k$ may display too little information about the table. Fortunately, the BRS algorithm is incremental in nature. That is, in order to find the best rule list of size $k+1$, it first finds the best rule-list of size $k$, and then finds another rule to add to get a rule-list of size $k+1$. Thus instead of running the algorithm with a fixed value of $k$, it can start with an empty rule-list and keep adding rules to it, displaying new rules as they are found. This search for additional rules can stop when the user issues a new smart drill-down command to the system, or manually stops the search. Alternatively, we can set a time limit (of say $5$ seconds) and display as many rules as we can find within that time limit.

$W$ is the weight function that determines which rules are interesting. This is a function specified by the user as a black box. Specifying an arbitrary function can be hard, so instead we hardcode some common Weight functions and allow the user to choose one from a drop-down menu. In addition, the user can express interest or disinterest in certain columns by telling the system to favor or ignore those columns, via the user interface. The system internally adjusts the weight function by increasing or decreasing the weight given to rules instantiating that column.

The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time. Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time. We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling. We create a small random sample of tuples from the table, and run the BRS algorithm on it. Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output. To account for sampling error, we can set $m_w$ to $2x$, which works well in practice.
}

### 2015-12-25 TableSummarization/Paper/exp.tex
addressed most reviewer comments, brought back to length

A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules. We use this to perform a drill down on the `Age' column using our experimental prototype. The result is shown in Figure~\ref{fig:drilldownexample}. We can contrast it with Figure~\ref{fig:uiexample1}; the latter gives information about multiple columns at once and only displays high count values. Regular drill down on the other hand, serves a complementary purpose by focusing on detailed evaluation of a single column.

@@

\mrj{Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.}

@@

\subsubsection{Scaling properties of our algorithms}
\mrj{
The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$. That is, the runtime can be written as $a \times |T| + b \times minSS$ where $a$ and $b$ are constants. In the worst-case where we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk. That is, $a \times |T|$ is the time taken to make a single sequential scan over the table on disk. The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.}

\mrj{When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a). When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk). When we have a few million tuples, our total runtime is only a few seconds. But if the dataset consisted of billions of tuples, the process of reading the table to create a sample could itself take a very long time. To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).
}

>>>>

A regular drill down can be thought of as a special case of smart drill-down with the right weighting function and number of rules\papertext{ (see technical report for details)}\techreporttext{. Specifically, if we want to perform a regular drill down on a column $C$ that has $n$ distinct values, then we set the number of rules to be displayed ($k$) to $n$. The weighting function $W$ is set such that $W(r) = 1$ if rule $r$ has $C$ instantiated, and $0$ otherwise. This ensures that all displayed rules have $C$ instantiated (with no other column $C'$ instantiated unless there is a functional dependency from $C$ to $C'$), and that each displayed rule has a distinct value of $C$. This effectively gives us a regular drill-down on $C$}. We use this to perform a drill down on the `Age' column using our experimental prototype. The result is shown in Figure~\ref{fig:drilldownexample}. We can contrast it with Figure~\ref{fig:uiexample1}; the latter gives information about multiple columns at once and only displays high count values. Regular drill down on the other hand, serves a complementary purpose by focusing on detailed evaluation of a single column.

@@

Figure~\ref{fig:minSS_speed}(a) shows that sampling gives us noticeable time savings. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.

@@

\subsubsection{Scaling properties of our algorithms}
The computation time for a smart drill-down is linear in both the table size $|T|$ and in parameter $minSS$. That is, the runtime can be written as $a \times |T| + b \times minSS$ where $a$ and $b$ are constants. In the worst-case where we cannot form a sample from main memory and need to re-create a sample, $a$ stands for the time taken to read data from hard disk. That is, $a \times |T|$ is the time taken to make a single sequential scan over the table on disk. The constant $b$ is bigger than $a$, because BRS makes multiple passes over the sample, while creating a sample only requires a single pass over the table.

When the $|T|$ is small, the runtme is dominated by the $b \times minSS$ term, as seen for the Marketing Dataset in Figure~\ref{fig:minSS_speed}(a). When $|T|$ is large relative to $minSS$, like for the Census Dataset, the runtime is dominated by the $a \times |T|$ term (this is when we need to create a fresh sample from hard disk). When we have a few million tuples, our total runtime is only a few seconds. But if the dataset consisted of billions of tuples, the process of reading the table to create a sample could itself take a very long time. To counteract this, we could preprocess the dataset by down-sampling it to only a million tuples, and perform the summarization on the million tuple sample (which also effectively summarizes the billion tuple table).

### 2015-12-25 TableSummarization/Paper/formal.tex
addressed most reviewer comments, brought back to length

Now, onto the second portion: we let $W$ denote a function that assigns a non-negative {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
As we will see, the weighting function does not depend on the specific
tuples in $\calD$, but could, as we will see later, 
depend on the number of $\star$s in $r$,
the schema of $\calD$, 
as well as the number of distinct values in each column of $\calD$.
A weighting function is said to be {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$; we focus
on monotonic weighting functions because we prefer 
rules that are more ``specific''
rather than those that are more ``general'' 
(thereby conveying less information).
We further describe our weighting functions in Section~\ref{sec:weighting}.

Thus, the total score for our list of rules is given by 
$$\text{Score}(R) = \sum_{r \in R} \underbrace{MCount(r, R)}_{\text{coverage of $r$ in $\calD$}} \times \underbrace{W(r)}_{\text{weight of $r$}}$$ 
Our goal is to choose the rule-list maximizing 
total score.

@@

\stitle{Formal Problem:} We now formally define our problem:
\begin{problem}\label{prob:optimal-subrule-list}
Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r) \times MCount(r,R)$$
for one of the following smart drill down operations:
\begin{denselist}
\item $[$Rule drill down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item $[$Star drill down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{denselist}
\end{problem}
\noindent Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user. We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead. We discuss how to modify our algorithms to use $Sum$ instead of $Count$ 
\papertext{in our technical report~\cite{tr}}.
\techreporttext{in the `Extensions' section}

@@

\techreporttext{
\begin{table}
\centering
\scriptsize
\begin{tabular}{ | l | c | }
 \hline Rule-MCount list & Score \\ \hline
  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
\end{tabular}
\vspace{-10pt}
\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
\vspace{-15pt}
\end{table}
}

\subsection{Smart drill down Operations}
\label{sec:interface}
When the user starts using a system equipped with
the smart drill down operator, they first see a table with a single trivial rule as shown in Table~\ref{table:introexample0}. At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill down' on the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}. Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on a $\star$, say in column $c$ of rule $r$. This will also cause rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. \techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product.} This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise. Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. \techreporttext{ For example, clicking on the walmart rule in Table~\ref{table:introexample2} would take the user back to Table~\ref{table:introexample}. This operation is equivalent to a traditional
roll up, but for smart drill downs instead of traditional drill downs.}

\techreporttext{
\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\vspace{-10pt}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\vspace{-10pt}
\end{table}
}

>>>>

Now, onto the second portion: we let $W$ denote a function that assigns a non-negative {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
As we will see, the weighting function does not depend on the specific
tuples in $\calD$, but could 
depend on the number of $\star$s in $r$,
the schema of $\calD$, 
as well as the number of distinct values in each column of $\calD$.
A weighting function is said to be {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$; we focus
on monotonic weighting functions because we prefer 
rules that are more ``specific''
rather than those that are more ``general'' 
(thereby conveying less information).
We further describe our weighting functions in Section~\ref{sec:weighting}.

Thus, the total score for our list of rules is given by 
$$\text{Score}(R) = \sum_{r \in R} \underbrace{MCount(r, R)}_{\text{coverage of $r$ in $\calD$}} \times \underbrace{W(r)}_{\text{weight of $r$}}$$ 
Our goal is to choose the rule-list of a given length that maximizes 
total score.

@@

\stitle{Formal Problem:} We now formally define our problem:
\begin{problem}\label{prob:optimal-subrule-list}
Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r) \times MCount(r,R)$$
for one of the following smart drill down operations:
\begin{denselist}
\item $[$Rule drill down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item $[$Star drill down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{denselist}
\end{problem}
\noindent Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user. We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead. We discuss how to modify our algorithms to use $Sum$ instead of $Count$ 
\papertext{in our technical report~\cite{tr}}.
\techreporttext{in the `Extensions' section.}

@@

\techreporttext{
\begin{table}
\centering
\scriptsize
\begin{tabular}{ | l | c | }
 \hline Rule-MCount list & Score \\ \hline
  ($a$, $b_1$)-$100$, ($a$, $\star$)-$900$ & $1100$ \\
  ($a$, $b_1$)-$100$, ($a$, $b_2$)-$300$, ($a$, $\star$)-$600$ & $1400$  \\
  ($a$, $b_1$)-$100$, ($a$, $b_3$)-$400$, ($a$, $\star$)-$500$ & $1500$ \\ \hline
\end{tabular}
\caption{Example of Rule-based scoring with score equal to rule size \label{table:sizescoringexample}}
\end{table}
}

\subsection{Smart drill down Operations}
\label{sec:interface}
When the user starts using a system equipped with
the smart drill down operator, they first see a table with a single trivial rule as shown in Table~\ref{table:introexample0}. At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill down' on the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user.\techreporttext{ As an example, clicking on the trivial rule of Table~\ref{table:introexample0} would display Table~\ref{table:introexample}. Clicking further on the third rule in the expanded rule-list would display Table~\ref{table:introexample2}.} The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on a $\star$, say in column $c$ of rule $r$. This will also cause rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. \techreporttext{ For instance, if the user clicks on the $\star$ in the product column of the walmart rule, they will see Table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product. This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.} Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. \techreporttext{ For example, clicking on the walmart rule in Table~\ref{table:introexample2} would take the user back to Table~\ref{table:introexample}. This operation is equivalent to a traditional
roll up, but for smart drill downs instead of traditional drill downs.}

\techreporttext{
\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\end{table}
}

### 2015-12-25 TableSummarization/Paper/intro.tex
addressed most reviewer comments, brought back to length

In addition to presenting this new operator called smart drill down, we also present novel sampling techniques to compute the results for this
operator {\em in an interactive fashion} on increasingly larger databases. 
Unlike the traditional OLAP setting, these computations 
require no pre-materialization, and can be implemented 
within or on top of any relational database system.

@@

When the analyst clicks on a rule $r$, smart drill down
expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
There are three factors that make a rule set interesting.
One is if it contains rules with high Count (or total sales) fields,
since the larger the count, the more tuples are summarized.
A second factor is if the rules have high weight (number of non-$\star$ attributes).
For instance, the rule (Walmart, cookies, AK-1, $200$, $3$)
seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
since the former tells us the high sales are concentrated in a single region.
A third desirability factor is diversity:
For example, if we already have the rule (Walmart, $\star$, $\star$, $1000$, $1$)
in our set, we would rather have the rule (Target, bicycles, $\star$, $200$, $2$)
than (Walmart, bicycles, $\star$, $200$, $2$) since the former rule
describes tuples that are not described by the first rule.

@@

\end{example}

@@

Our work on smart drill down is related
to table summarization and anomaly
detection~\cite{Sarawagi:2001:UMA:767141.767148,
Sarawagi00user-adaptiveexploration,
Sarawagi98discovery-drivenexploration,
DBLP:journals/pvldb/GebalyAGKS14}.
These papers mostly focus on
giving the most ``surprising'' information to the user, i.e., information
that would minimize the Kullback-Liebler(KL) divergence between the
resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
values occur together in an unexpectedly small number of tuples, that
set of values may be displayed to the user. In contrast, our algorithm
focuses on rules with high counts, covering as
much of the table as possible.
Furthermore, our summarization is couched in
an interactive environment, where the analyst
directs the drill down and can tailor the optimization criteria.
Nevertheless, one can envision extending traditional
and smart drill down to provide an additional option of anomaly detection.

@@

\noindent {\bf Contributions.} Our chief contribution in this paper is the {\em smart drill down} interaction operator,
an extension of traditional drill down, aimed at allowing analysts to zoom into the
more ``interesting'' parts of a dataset. In addition to this operator,
we develop techniques to support this operator on increasingly larger datasets:
\squishlist
\item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. 
\item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
maintaining samples in memory to support smart drill down. 
We show that optimal identification of samples is once again {\sc NP-Hard}, and 
we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
\squishend
We have developed a {\em fully functional and usable prototype tool} that supports the smart drill-down operator that was demonstrated at VLDB this year~\cite{DBLP:journals/pvldb/JoglekarGP15}. From this point on, when we provide result snippets, these will be screenshots from our prototype tool. Our prototype tool also supports traditional drill-down: smart drill-down 
can be viewed as a generalization of traditional drill-down (with the weighting function set appropriately). In Section~\ref{sec:exp_qualitative}, we compare smart drill-down with traditional drill-down and show that smart drill-down returns considerably better results.

>>>>

In addition to presenting the new smart drill down operator, we present novel sampling techniques to compute the results for this
operator {\em in an interactive fashion} on increasingly larger databases. 
Unlike the traditional OLAP setting, these computations 
require no pre-materialization, and can be implemented 
within or on top of any relational database system.

@@

When the analyst clicks on a rule $r$, smart drill down
expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
There are three factors that make a rule set interesting.
One is if it contains rules with high Count (or total sales) fields,
since the larger the count, the more tuples are summarized.
A second factor is if the rules have high weight (number of non-$\star$ attributes).
For instance, the rule (Walmart, cookies, AK-1, $200$, $3$)
seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
since the former tells us the high sales are concentrated in a single region.
A third desirability factor is diversity:
For example, if we already have the rule (Walmart, $\star$, $\star$, $1000$, $1$)
in our set, we would rather have the rule (Target, bicycles, $\star$, $200$, $2$)
than (Walmart, bicycles, $\star$, $200$, $2$) since the former rule
describes tuples that are not described by the first rule.
\end{example}

@@

@@

Our work on smart drill down is related
to table summarization and anomaly
detection~\cite{Sarawagi:2001:UMA:767141.767148,
Sarawagi00user-adaptiveexploration,
Sarawagi98discovery-drivenexploration,
DBLP:journals/pvldb/GebalyAGKS14}.
These papers mostly focus on
giving the most ``surprising'' information to the user, i.e., information
that would minimize the Kullback-Liebler(KL) divergence between the
resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
values occur together in an unexpectedly small number of tuples, that
set of values may be displayed to the user. In contrast, our algorithm
focuses on rules with high counts, covering as
much of the table as possible. Thus our work can be thought of as complementary to anomaly detection.
Furthermore, our summarization is couched in
an interactive environment, where the analyst
directs the drill down and can tailor the optimization criteria.

@@

\noindent {\bf Contributions.} Our chief contribution in this paper is the {\em smart drill down} interaction operator,
an extension of traditional drill down, aimed at allowing analysts to zoom into the
more ``interesting'' parts of a dataset. In addition to this operator,
we develop techniques to support this operator on increasingly larger datasets:
\squishlist
\item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. 
\item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
maintaining samples in memory to support smart drill down. 
We show that optimal identification of samples is once again {\sc NP-Hard}, and 
we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
\squishend
We have developed a {\em fully functional and usable prototype tool} that supports the smart drill-down operator that was demonstrated at VLDB 2015~\cite{DBLP:journals/pvldb/JoglekarGP15}. From this point on, when we provide result snippets, these will be screenshots from our prototype tool. Our prototype tool also supports traditional drill-down: smart drill-down 
can be viewed as a generalization of traditional drill-down (with the weighting function set appropriately). In Section~\ref{sec:exp_qualitative}, we compare smart drill-down with traditional drill-down and show that smart drill-down returns considerably better results.

### 2015-12-25 TableSummarization/Paper/sampling.tex
addressed most reviewer comments, brought back to length

We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user). \mrj{$M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.}
The parameter $minSS$ is the minimum number of sample tuples needed
such that we can use the sample in memory 
instead of having to resort to the entire table stored
on hard disk. 
This parameter determines how accurate our count estimates will be, and also
how quickly we can return results to the user. \mrj{The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.}

>>>>

We are given a memory capacity or budget $M$, and a minimum sample size $minSS$ (both specified by the user). $M$ can simply be set to the actual available memory, while $minSS$ can easily be computed as a function of the desired estimation error.
The parameter $minSS$ is the minimum number of sample tuples needed
such that we can use the sample in memory 
instead of having to resort to the entire table stored
on hard disk. 
This parameter determines how accurate our count estimates will be, and also
how quickly we can return results to the user. The count estimate error is proportional to $\sqrt{\frac{1}{minSS}}$ while runtime is proportional to $minSS$.
